---
title: "Demo. asignacion-1"
output:
  word_document: default
  html_document: default
date: "2023-05-04"
---

```{r}
load("~/Bases/pregunta1.a.Rdata")
```

Grafico
```{r}
library(ggplot2)
ggplot(base, aes(base$Edad,base$total)) + geom_point() + geom_smooth(span=0.3, se=FALSE)+ggtitle("grafico 1")
```

```{r}

ggplot(base, aes(base$Edad,base$total)) + geom_line()+ggtitle("grafico 2")
```

```{r}
ggplot(base,aes(base$Edad,base$total)) + geom_point() + geom_smooth(span=0.9, se=FALSE) + ggtitle("grafico 3")
```

```{r}
x=loess(base$total~base$Edad,span = 0.25)
nueva=predict(x)
cbind(nueva)
```

```{r}
nueva=cbind(nueva)
base$edad_corregida=nueva[,1]
colnames(base)=c("Hombre","mujer","total","Edad","total.corregida")
View(base)
```

```{r}
ggplot(base, aes(base$Edad,base$total.corregida)) + geom_line()+ggtitle("grafico 3 total edad corregida")
```
```{r}
save(base,file = "base.corregida.Rdata")
```

nueva base con las poblaciones corregidas
```{r}
load("~/Bases/base.corregida.Rdata")
View(base)
```


```{r}
base$Edad.corregida= ceiling(base$Edad.corregida)
```


library(openxlsx); write.xlsx(base, "basecorregida.xlsx", sheetName = "basecorregida")


# Piramides
Costa rica

base=subset(Costa_Rica,Costa_Rica$Year=="2021")
base=subset(base,base$Area=="Total")
base=subset(base,base$Sex=="Male"|base$Sex=="Female")
View(base)

save(base,file="CR.Rdata")



load("CR.Rdata")
base=cbind(base$Age,base$Sex,base$Value)
colnames(base)=c("edad","sexo","total")
base=as.data.frame(base)
baseh=subset(base,base$sexo=="Male") ;basem=subset(base,base$sexo=="Female")
base=c(baseh,basem)
View(base)
base=base[,-c(2,4,5)]
colnames(base)=c("edad","ph","pm")
save(base,file = "piramideCR.Rdata")


```{r}
load("piramideCR.Rdata")
library(pyramid)
base$ph=as.numeric(base$ph)
base$pm=as.numeric(base$pm)
n=sum(base$ph)+sum(base$pm)
base$php=(base$ph/n)*100
base$pmp=(base$pm/n)*100

pyramids(base$php,base$pmp,Center=base$edad,
       Llab="Hombres",Rlab="Mujeres",
       main="Piramide poblacional Costa Rica 2022",Clab="edad",
       Laxis=c(0,0.25,0.5,0.85,1),Cstep=5,Cgap=0.2)
```

Afghanistan

base=subset(Afghanistan,Afghanistan$Year=="1979")
base=subset(base,base$Area=="Total")
base=subset(base,base$Sex=="Male"|base$Sex=="Female")
View(base)



save(base,file="Afghanistan.Rdata")



load("Afghanistan.Rdata")
base=cbind(base$Age,base$Sex,base$Value)
colnames(base)=c("edad","sexo","total")
base=as.data.frame(base)
baseh=subset(base,base$sexo=="Male") ;basem=subset(base,base$sexo=="Female")
base=c(baseh,basem)
base=as.data.frame(base)
View(base)
base=base[,c(1,3,6)]
colnames(base)=c("edad","ph","pm")
save(base,file = "piramideaAFGHA.Rdata")


```{r}
load("piramideaAFGHA.Rdata")
library(pyramid)
base$ph=as.numeric(base$ph)
base$pm=as.numeric(base$pm)
n=sum(base$ph)+sum(base$pm)
base$php=(base$ph/n)*100
base$pmp=(base$pm/n)*100

pyramids(base$php,base$pmp,Center=base$edad,
         Llab="Hombres",Rlab="Mujeres",
         main="Piramide poblacional Afghanistan 1979",Clab="edad",
         Laxis=c(0,1,2,3,4,5.5,7,8),Cstep=2,Cgap=0.3)
```

Japon

base=subset(Japan,Japan$Year=="2021")
base=subset(base,base$Area=="Total")
base=subset(base,base$Sex=="Male"|base$Sex=="Female")
View(base)



save(base,file="Japan.Rdata")




load("Japan.Rdata")
base=cbind(base$Age,base$Sex,base$Value)
colnames(base)=c("edad","sexo","total")
base=as.data.frame(base)
baseh=subset(base,base$sexo=="Male") ;basem=subset(base,base$sexo=="Female")
base=c(baseh,basem)
base=as.data.frame(base)
View(base)
base=base[,c(1,3,6)]
colnames(base)=c("edad","ph","pm")
#basequinquenal=base[c(1,3,9,15,21,27,33,39,45,51,57,63,69,75,81,87,93,99,105,111,117,123),]
base=base[-c(1,3,9,15,21,27,33,39,45,51,57,63,69,75,81,87,93,99,105,111,117,123),]
save(base,file = "piramideJapan.Rdata")


Piramide japan
```{r}
load("piramideJapan.Rdata")
library(pyramid)
base$ph=as.numeric(base$ph)
base$pm=as.numeric(base$pm)
n=sum(base$ph)+sum(base$pm)
base$php=(base$ph/n)*100
base$pmp=(base$pm/n)*100

pyramids(base$php,base$pmp,Center=base$edad,
         Llab="Hombres",Rlab="Mujeres",
         main="Piramide poblacional japan 2021",Clab="edad",
         Laxis=c(0,0.3,0.5,0.6,0.75,0.9),Cstep=8,Cgap=0.1)
```

Bulgaria

base=subset(Bulgaria,Bulgaria$Year=="2021")
base=subset(base,base$Area=="Total")
base=subset(base,base$Sex=="Male"|base$Sex=="Female")
View(base)



save(base,file="Bulgaria.Rdata")




load("Bulgaria.Rdata")
base=cbind(base$Age,base$Sex,base$Value)
colnames(base)=c("edad","sexo","total")
base=as.data.frame(base)
baseh=subset(base,base$sexo=="Male") ;basem=subset(base,base$sexo=="Female")
base=c(baseh,basem)
base=as.data.frame(base)
View(base)
base=base[,c(1,3,6)]
colnames(base)=c("edad","ph","pm")
#basequinquenal=base[c(1,3,5,10,16,22,28,34,40,46,52,58,64,70,76,82,88,94,100,106,112,118,124,125,131),]
base=base[-c(1,3,5,10,16,22,28,34,40,46,52,58,64,70,76,82,88,94,100,106,112,118,124,125,131),]
save(base,file = "piramideBulgaria.Rdata")


Piramide bulgaria
```{r}
load("piramideBulgaria.Rdata")
library(pyramid)
base$ph=as.numeric(base$ph)
base$pm=as.numeric(base$pm)
n=sum(base$ph)+sum(base$pm)
base$php=(base$ph/n)*100
base$pmp=(base$pm/n)*100

pyramids(base$php,base$pmp,Center=base$edad,
         Llab="Hombres",Rlab="Mujeres",
         main="Piramide poblacional Bulgaria 2021",Clab="edad",
         Laxis=c(0,0.25,0.5,0.6,0.75,0.9),Cstep=8,Cgap=0.1)
```

Guatemala

base=subset(Guatemala,Guatemala$Year=="2020")
base=subset(base,base$Area=="Total")
base=subset(base,base$Sex=="Male"|base$Sex=="Female")
View(base)



save(base,file="Guatemala.Rdata")



load("Guatemala.Rdata")
base=cbind(base$Age,base$Sex,base$Value)
colnames(base)=c("edad","sexo","total")
base=as.data.frame(base)
baseh=subset(base,base$sexo=="Male") ;basem=subset(base,base$sexo=="Female")
base=c(baseh,basem)
base=as.data.frame(base)
View(base)
base=base[,c(1,3,6)]
colnames(base)=c("edad","ph","pm")
basequinquenal=base[c(1,3,9,15,21,27,33,39,45,51,57,63,69,75,81,87,93,99,105,111,117,122),]
base=base[-c(1,3,9,15,21,27,33,39,45,51,57,63,69,75,81,87,93,99,105,111,117,122),]
save(base,file = "piramideGuatemala.Rdata")


Piramide Guatemala
```{r}
load( "piramideGuatemala.Rdata")
library(pyramid)
base$ph=as.numeric(base$ph)
base$pm=as.numeric(base$pm)
n=sum(base$ph)+sum(base$pm)
base$php=(base$ph/n)*100
base$pmp=(base$pm/n)*100

pyramids(base$php,base$pmp,Center=base$edad,
         Llab="Hombres",Rlab="Mujeres",
         main="Piramide poblacional Guatemala 2020",Clab="edad",
         Laxis=c(0,0.3,0.4,0.6,0.8,1,1.2),Cstep=5,Cgap=0.1)
```
##costarica 1963


base=costa.rica.93
colnames(base)=c("edad","hombre","mujer","total")
base=base[-1,]



save(base,file="costa.rica.93.Rdata")



load("costa.rica.93.Rdata")
base=as.data.frame(base)
base=base[,c(1,2,3)]
colnames(base)=c("edad","ph","pm")
base=base[-100,]
save(base,file = "costa.rica.93.Rdata")


```{r}
load("costa.rica.93.Rdata")
library(pyramid)
base$ph=as.numeric(base$ph)
base$pm=as.numeric(base$pm)
n=sum(base$ph)+sum(base$pm)
base$php=(base$ph/n)*100
base$pmp=(base$pm/n)*100

pyramids(base$php,base$pmp,Center=base$edad,
         Llab="Hombres",Rlab="Mujeres",
         main="Piramide poblacional Costa Rica 1993",Clab="edad",
         Laxis=c(0.001,0.5,0.7,1,2),Cstep=5,Cgap=0.2)
```



